---
title: "Identifying Trends in H2B Visa Certification"
author: "24-7"
date: "5-4-18"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r - install_packages, include=FALSE}
library(tidyverse)
library(readxl)
library(stringr)
library(maps)
library(Metrics)
library(shiny)
```

```{r-seed}
set.seed(1234)
```


```{r join-data, echo = FALSE, message=FALSE}
h2b2008 <- read_csv("data/H-2B_FY2008.csv") %>%
  mutate(year = 2008)
h2b2009 <- read_csv("data/H-2B_FY2009.csv") %>%
  mutate(year = 2009)
h2b2010 <- read_csv("data/H-2B_FY2010.csv") %>%
  mutate(year = 2010)
h2b2011 <- read_csv("data/H-2B_FY2011.csv") %>%
  mutate(year = 2011)
h2b2012 <- read_csv("data/H-2B_FY2012.csv") %>%
  mutate(year = 2012)

h2b <- h2b2008 %>%
  full_join(h2b2009) %>%
  full_join(h2b2010)%>%
  full_join(h2b2011)%>%
  full_join(h2b2012)

h2b <- h2b %>%
  select(year, CASE_NO, DECISION_DATE, NPC_SUBMITTED_DATE, CASE_STATUS, ALIEN_WORK_STATE, CERTIFICATION_BEGIN_DATE, CERTIFICATION_END_DATE, EMPLOYER_CITY, EMPLOYER_STATE, EMPLOYER_POSTAL_CODE, AGENT_ATTORNEY_CITY, AGENT_ATTORNEY_STATE, JOB_TITLE, NBR_WORKERS_CERTIFIED, PREVAILING_WAGE, PW_UNIT_OF_PAY, BASIC_RATE_OF_PAY, BASIC_UNIT_OF_PAY, DOT_OCCUPATIONAL_CODE, NBR_WORKERS_REQUESTED, DOT_NAME, SOC_CODE, SOC_NAME)
```

### Introduction
An H2-B visa is a temporary work visa for foreign workers with a job offer for seasonal, non-agricultural work in the United States. Employers from the United States must first submit temporary labor certification applications to the Department of Labor to allow H-2B workers to apply for visas. (Source: U.S. Citizenship and Immigration Services) The H2-B visa is especially important for industries with seasonal variations, such as landscaping and hospitality (Source: http://thehill.com/blogs/congress-blog/labor/269298-the-nation-needs-the-h2b-program). Small businesses rely on these temporary workers to keep their business running during peak times. However, not all applications for this visa are approved. Currently, H2-B visas are capped at 66,000 per fiscal year. (Source: https://www.uscis.gov/working-united-states/temporary-workers/h-2b-temporary-non-agricultural-workers). Through analyzing data provided by the Department of Labor's Office of Foreign Labor Certification, this project aims to find out what factors make applicants more likely to get successfully obtain H2-B visas. Specifically, this project looks at how job category, wage, season, region, and number of workers affects the likelihood of getting an H2-B visa application approved.

In this project, we produced visualizations to present the data in an accessible, digestible format. We created a visualization for each of the factors we are considering. Furthermore, we created a regression model to further analyze how the variables affect the likelihood of a case being approved. Finally, we created a shiny app so users can see the relationships between various factors. 



```{r - uncapitalize_variables}
h2b <- h2b %>%
  mutate(case_no = CASE_NO, 
         decision_date = DECISION_DATE,
         npc_submitted_date = NPC_SUBMITTED_DATE,
         case_status = CASE_STATUS, 
         alien_work_state = ALIEN_WORK_STATE, 
         certification_begin_date = CERTIFICATION_BEGIN_DATE, 
         certification_end_date = CERTIFICATION_END_DATE, 
         employer_city = EMPLOYER_CITY, 
         employer_state = EMPLOYER_STATE, 
         employer_postal_code = EMPLOYER_POSTAL_CODE, 
         agent_attorney_city = AGENT_ATTORNEY_CITY, 
         agent_attorney_state = AGENT_ATTORNEY_STATE, 
         job_title = JOB_TITLE, 
         nbr_workers_certified = NBR_WORKERS_CERTIFIED, 
         prevailing_wage = PREVAILING_WAGE, 
         pw_unit_of_pay = PW_UNIT_OF_PAY, 
         basic_rate_of_pay = BASIC_RATE_OF_PAY, 
         basic_unit_of_pay = BASIC_UNIT_OF_PAY, 
         dot_occupational_code = DOT_OCCUPATIONAL_CODE, 
         nbur_workers_requested =NBR_WORKERS_REQUESTED, 
         dot_name = DOT_NAME, 
         soc_code = SOC_CODE, 
         soc_name = SOC_NAME)

h2b <- h2b %>%
  select(year, case_no, decision_date, npc_submitted_date, case_status, alien_work_state, certification_begin_date, certification_end_date, employer_city, employer_state, employer_postal_code, agent_attorney_city, agent_attorney_state, job_title, nbr_workers_certified, prevailing_wage, pw_unit_of_pay, basic_rate_of_pay, basic_unit_of_pay, dot_occupational_code, nbur_workers_requested, dot_name, soc_code, soc_name)

```


```{r - muate_benefits}

h2b <- h2b %>%
  mutate(wage_diff = case_when(
  prevailing_wage > basic_rate_of_pay ~ "prevail_higher",
  prevailing_wage == basic_rate_of_pay ~ "same",
  prevailing_wage < basic_rate_of_pay ~ "basic_higher"
  ))
  

```


```{r - mutate_season}
h2b <- h2b %>%
  mutate(decision_season = case_when(
    str_sub(decision_date,1,2) == "1/" ~ "winter",
    str_sub(decision_date,1,2) == "2/" ~ "winter",
    str_sub(decision_date,1,2) == "3/" ~ "spring",
    str_sub(decision_date,1,2) == "4/" ~ "spring",
    str_sub(decision_date,1,2) == "5/" ~ "spring",
    str_sub(decision_date,1,2) == "6/" ~ "summer",
    str_sub(decision_date,1,2) == "7/" ~ "summer",
    str_sub(decision_date,1,2) == "8/" ~ "summer",
    str_sub(decision_date,1,2) == "9/" ~ "fall",
    str_sub(decision_date,1,3) == "10/" ~ "fall",
    str_sub(decision_date,1,3) == "11/" ~ "fall",
    str_sub(decision_date,1,3) == "12/" ~ "winter",
    str_sub(decision_date,3,5) == "Jan" ~ "winter",
    str_sub(decision_date,4,6) == "Jan" ~ "winter",
    str_sub(decision_date,3,5) == "Feb" ~ "winter",
    str_sub(decision_date,4,6) == "Feb" ~ "winter",
    str_sub(decision_date,3,5) == "Mar" ~ "spring",
    str_sub(decision_date,4,6) == "Mar" ~ "spring",
    str_sub(decision_date,3,5) == "Apr" ~ "spring",
    str_sub(decision_date,4,6) == "Apr" ~ "spring",
    str_sub(decision_date,3,5) == "May" ~ "spring",
    str_sub(decision_date,4,6) == "May" ~ "spring",
    str_sub(decision_date,3,5) == "Jun" ~ "summer",
    str_sub(decision_date,4,6) == "Jun" ~ "summer",
    str_sub(decision_date,3,5) == "Jul" ~ "summer",
    str_sub(decision_date,4,6) == "Jul" ~ "summer",
    str_sub(decision_date,3,5) == "Aug" ~ "summer",
    str_sub(decision_date,4,6) == "Aug" ~ "summer",
    str_sub(decision_date,3,5) == "Sep" ~ "fall",
    str_sub(decision_date,4,6) == "Sep" ~ "fall",
    str_sub(decision_date,3,5) == "Oct" ~ "fall",
    str_sub(decision_date,4,6) == "Oct" ~ "fall",
    str_sub(decision_date,3,5) == "Nov" ~ "fall",
    str_sub(decision_date,4,6) == "Nov" ~ "fall",
    str_sub(decision_date,3,5) == "Dec" ~ "winter",
    str_sub(decision_date,4,6) == "Dec" ~ "winter",
  ))

h2b <- h2b %>%
  mutate(submitted_season = case_when(
    str_sub(npc_submitted_date,1,2) == "1/" ~ "winter",
    str_sub(npc_submitted_date,1,2) == "2/" ~ "winter",
    str_sub(npc_submitted_date,1,2) == "3/" ~ "spring",
    str_sub(npc_submitted_date,1,2) == "4/" ~ "spring",
    str_sub(npc_submitted_date,1,2) == "5/" ~ "spring",
    str_sub(npc_submitted_date,1,2) == "6/" ~ "summer",
    str_sub(npc_submitted_date,1,2) == "7/" ~ "summer",
    str_sub(npc_submitted_date,1,2) == "8/" ~ "summer",
    str_sub(npc_submitted_date,1,2) == "9/" ~ "fall",
    str_sub(npc_submitted_date,1,3) == "10/" ~ "fall",
    str_sub(npc_submitted_date,1,3) == "11/" ~ "fall",
    str_sub(npc_submitted_date,1,3) == "12/" ~ "winter",
    str_sub(npc_submitted_date,3,5) == "Jan" ~ "winter",
    str_sub(npc_submitted_date,4,6) == "Jan" ~ "winter",
    str_sub(npc_submitted_date,3,5) == "Feb" ~ "winter",
    str_sub(npc_submitted_date,4,6) == "Feb" ~ "winter",
    str_sub(npc_submitted_date,3,5) == "Mar" ~ "spring",
    str_sub(npc_submitted_date,4,6) == "Mar" ~ "spring",
    str_sub(npc_submitted_date,3,5) == "Apr" ~ "spring",
    str_sub(npc_submitted_date,4,6) == "Apr" ~ "spring",
    str_sub(npc_submitted_date,3,5) == "May" ~ "spring",
    str_sub(npc_submitted_date,4,6) == "May" ~ "spring",
    str_sub(npc_submitted_date,3,5) == "Jun" ~ "summer",
    str_sub(npc_submitted_date,4,6) == "Jun" ~ "summer",
    str_sub(npc_submitted_date,3,5) == "Jul" ~ "summer",
    str_sub(npc_submitted_date,4,6) == "Jul" ~ "summer",
    str_sub(npc_submitted_date,3,5) == "Aug" ~ "summer",
    str_sub(npc_submitted_date,4,6) == "Aug" ~ "summer",
    str_sub(npc_submitted_date,3,5) == "Sep" ~ "fall",
    str_sub(npc_submitted_date,4,6) == "Sep" ~ "fall",
    str_sub(npc_submitted_date,3,5) == "Oct" ~ "fall",
    str_sub(npc_submitted_date,4,6) == "Oct" ~ "fall",
    str_sub(npc_submitted_date,3,5) == "Nov" ~ "fall",
    str_sub(npc_submitted_date,4,6) == "Nov" ~ "fall",
    str_sub(npc_submitted_date,3,5) == "Dec" ~ "winter",
    str_sub(npc_submitted_date,4,6) == "Dec" ~ "winter",
  ))

h2b <- h2b %>%
  mutate(start_season = case_when(
    str_sub(certification_begin_date,1,2) == "1/" ~ "winter",
    str_sub(certification_begin_date,1,2) == "2/" ~ "winter",
    str_sub(certification_begin_date,1,2) == "3/" ~ "spring",
    str_sub(certification_begin_date,1,2) == "4/" ~ "spring",
    str_sub(certification_begin_date,1,2) == "5/" ~ "spring",
    str_sub(certification_begin_date,1,2) == "6/" ~ "summer",
    str_sub(certification_begin_date,1,2) == "7/" ~ "summer",
    str_sub(certification_begin_date,1,2) == "8/" ~ "summer",
    str_sub(certification_begin_date,1,2) == "9/" ~ "fall",
    str_sub(certification_begin_date,1,3) == "10/" ~ "fall",
    str_sub(certification_begin_date,1,3) == "11/" ~ "fall",
    str_sub(certification_begin_date,1,3) == "12/" ~ "winter",
    str_sub(certification_begin_date,3,5) == "Jan" ~ "winter",
    str_sub(certification_begin_date,4,6) == "Jan" ~ "winter",
    str_sub(certification_begin_date,3,5) == "Feb" ~ "winter",
    str_sub(certification_begin_date,4,6) == "Feb" ~ "winter",
    str_sub(certification_begin_date,3,5) == "Mar" ~ "spring",
    str_sub(certification_begin_date,4,6) == "Mar" ~ "spring",
    str_sub(certification_begin_date,3,5) == "Apr" ~ "spring",
    str_sub(certification_begin_date,4,6) == "Apr" ~ "spring",
    str_sub(certification_begin_date,3,5) == "May" ~ "spring",
    str_sub(certification_begin_date,4,6) == "May" ~ "spring",
    str_sub(certification_begin_date,3,5) == "Jun" ~ "summer",
    str_sub(certification_begin_date,4,6) == "Jun" ~ "summer",
    str_sub(certification_begin_date,3,5) == "Jul" ~ "summer",
    str_sub(certification_begin_date,4,6) == "Jul" ~ "summer",
    str_sub(certification_begin_date,3,5) == "Aug" ~ "summer",
    str_sub(certification_begin_date,4,6) == "Aug" ~ "summer",
    str_sub(certification_begin_date,3,5) == "Sep" ~ "fall",
    str_sub(certification_begin_date,4,6) == "Sep" ~ "fall",
    str_sub(certification_begin_date,3,5) == "Oct" ~ "fall",
    str_sub(certification_begin_date,4,6) == "Oct" ~ "fall",
    str_sub(certification_begin_date,3,5) == "Nov" ~ "fall",
    str_sub(certification_begin_date,4,6) == "Nov" ~ "fall",
    str_sub(certification_begin_date,3,5) == "Dec" ~ "winter",
    str_sub(certification_begin_date,4,6) == "Dec" ~ "winter",
  ))
```


```{r-mutate_region}
h2b <- h2b %>%
  mutate(region = case_when(
    employer_state == "WA" | employer_state == "OR" | employer_state == "CA" |  employer_state == "ID" | employer_state == "MT" | employer_state == "WY" | employer_state == "NV" | employer_state == "UT" | employer_state == "CO" | employer_state == "AZ" | employer_state == "NM" | employer_state == "AK" | employer_state == "HI" ~ "West",
    employer_state == "ND" | employer_state == "SD" | employer_state == "NE" | employer_state == "KS" | employer_state == "MN" | employer_state == "IA" | employer_state == "MO" | employer_state == "WI" | employer_state == "IL" | employer_state == "MI" | employer_state == "IN" | employer_state == "OH" ~ "Midwest",
    employer_state == "TX" | employer_state == "OK" | employer_state == "AR" | employer_state == "LA" | employer_state == "MS" | employer_state == "AL" | employer_state == "TN" | employer_state == "KY" | employer_state == "WV" | employer_state == "MD" | employer_state == "DE" | employer_state == "VA" | employer_state == "NC" | employer_state == "SC" | employer_state == "GA" | employer_state == "FL" ~ "South",
    employer_state == "ME" | employer_state == "NH" | employer_state == "VT" | employer_state == "MA" | employer_state == "CT" | employer_state == "RI" | employer_state == "NJ" | employer_state == "NY" | employer_state == "PA" ~ "Northeast"
  ))
```

```{r-mutate_job_categories}
h2b <- h2b %>%
  mutate(occupation_category = case_when(
    str_sub(dot_occupational_code,1,1) == "0" | str_sub(dot_occupational_code,1,1) == "1" ~ "professional, technical, and managerial",
    str_sub(dot_occupational_code,1,1) == "2" ~ "clerical and sales",
    str_sub(dot_occupational_code,1,1) == "3" ~ "service",
    str_sub(dot_occupational_code,1,1) == "4" ~ "agricultural, fishery, forestry, and related",
    str_sub(dot_occupational_code,1,1) == "5" ~ "processing",
    str_sub(dot_occupational_code,1,1) == "6" ~ "machine trades",
    str_sub(dot_occupational_code,1,1) == "7" ~ "benchwork",
    str_sub(dot_occupational_code,1,1) == "8" ~ "structural work",
    str_sub(dot_occupational_code,1,1) == "9" ~ "miscellaneous"
  ))

```

```{r - map}
state_props <- h2b %>%
  filter(!is.na(nbur_workers_requested) & !(is.na(nbr_workers_certified))) %>%
  filter(employer_state %in% state.abb | employer_state == "DC") %>%
  filter(employer_state != "HI" & employer_state!="AK") %>%
  mutate(region = case_when(employer_state %in% state.abb ~ tolower(setNames(state.name, state.abb))[employer_state], employer_state == "DC" ~ "district of columbia")) %>%
  group_by(region) %>%
  summarise(prop = sum(nbr_workers_certified) / sum(nbur_workers_requested))

map <- merge(state_props, map_data("state"), sort = FALSE, by = "region")


```

```{r line-graph}
h2b %>% 
  filter(case_status %in% c("CERTIFIED", "PARTIAL CERTIFIED", "DENIED")) %>%
  group_by(year) %>%
  count(case_status) %>%
  ggplot(mapping = aes(x = year, y = n, color = case_status)) +
  geom_point() + 
  geom_line() +
  labs(title = "Decrease in number of H2B visas certified",
       subtitle = "From 2008 to 2012",
       x = "Year",
       y = "Number",
       color = "Application Result")
```

In 2008, Congress failed to renew the Save Our Small and Seasonal Bussinesses Act (SOSSBA). This reduced the cap for the total number of H2B Visas that could be approved every year. 

```{r-mutate-prop-certified}
h2b <- h2b %>%
  mutate(prop_workers_certified = nbr_workers_certified/nbur_workers_requested) 
```

```{r-backwards-selection}
mod <- lm(prop_workers_certified ~ occupation_category + prevailing_wage + submitted_season + region + nbur_workers_requested, data = h2b)

selected_mod <- step(mod, direction = "backward")

(selected_mod)
```

Using backwards selection, we discovered that the number of workers requested was not a significant factor that determined the proportion of workers certified. Since case status is a categorical variable, the model predicts the proportion of workers certified for each visa application (a numerical variable) instead. In terms of which industries are most likely to get the H2-B visa approved, the model indicates that agricultural, fishery, forestry, and related (which counts as one occupational category) and processing approve the largest proportion of cases. On the other hand, visas for workers in professional, technical, and managerial work (which counts as one occupational category) are least likely to be approved. Relative to processing, a visa for a worker in a professional, technical, or managerial position is 21.69 percent less likely to be approved. This makes sense because processing and agricultural jobs are not desirable to American workers, so there is a labor shortage in these sectors. Furthermore, professional jobs are more sought after by Americans, and they are positions that don’t lend themselves as well to temporary work. Therefore, these jobs are not likely to be approved for a H2-B temporary visa.

Additionally, visas submitted in the winter have the highest chance of certification, followed by fall, summer, and then spring.  Winter has a 10.37 higher percent chance of approval than spring. As for regional differences, visas submitted in the Northeast are mostly likely to be approved, followed by the South, the Midwest, and finally the West. Visas submitted in the Northeast is 5.18 percent more likely to be approved than the West. 


```{r map-wage}
state_wages <- h2b %>%
  filter(!is.na(basic_rate_of_pay) & !(is.na(basic_unit_of_pay))) %>%
  filter(employer_state %in% state.abb | employer_state == "DC") %>%
  filter(employer_state != "HI" & employer_state!="AK") %>%
  filter(basic_unit_of_pay == "HR") %>%
  mutate(region = case_when(employer_state %in% state.abb ~ tolower(setNames(state.name, state.abb))[employer_state], employer_state == "DC" ~ "district of columbia")) %>%
  group_by(region) %>%
  summarise(median_wage = median(basic_rate_of_pay))

map <- merge(state_wages, map, sort = FALSE, by = "region")
```

```{r map-occupation}
state_occ <- h2b %>%
  filter(!is.na(occupation_category)) %>%
  filter(employer_state %in% state.abb | employer_state == "DC") %>%
  filter(employer_state != "HI" & employer_state!="AK") %>%
  mutate(region = case_when(employer_state %in% state.abb ~ tolower(setNames(state.name, state.abb))[employer_state], employer_state == "DC" ~ "district of columbia")) %>% 
   group_by(region) %>%
   count(region, occupation_category) %>%
  slice(which.max(n))

map <- merge(state_occ, map, sort = FALSE, by = "region")

```

```{r map-season}
state_season <- h2b %>%
  filter(!is.na(start_season)) %>%
  filter(employer_state %in% state.abb | employer_state == "DC") %>%
  filter(employer_state != "HI" & employer_state!="AK") %>%
  mutate(region = case_when(employer_state %in% state.abb ~ tolower(setNames(state.name, state.abb))[employer_state], employer_state == "DC" ~ "district of columbia")) %>% 
   group_by(region) %>%
   count(region, start_season) %>%
  slice(which.max(n))

map <- merge(state_season, map, sort = FALSE, by = "region")
```

```{r map-total}
state_total <- h2b %>%
  filter(employer_state %in% state.abb | employer_state == "DC") %>%
  filter(employer_state != "HI" & employer_state!="AK") %>%
  mutate(region = case_when(employer_state %in% state.abb ~ tolower(setNames(state.name, state.abb))[employer_state], employer_state == "DC" ~ "district of columbia")) %>% 
   group_by(region) %>%
   summarise(n = n())

map <- merge(state_total, map, sort = FALSE, by = "region")
```


```{r shiny-map}
ui <- fluidPage(
      titlePanel("H2B visas by state"),
      sidebarLayout(
        sidebarPanel(
          selectInput(inputId = "fill", label = "Fill Map By",
          choices = c("Proportion approved" = "prop",
                      "Total applications" = "n", 
                      "Median hourly wage" = "median_wage",
                      "Most popular occupation" = "occupation_category",
                      "Most popular season" = "start_season"),
          selected = "Proportion approved")
        ),
       mainPanel(
         plotOutput(outputId = "map")
       )
      )
      
  )

server <- function(input, output) {
  output$map <- renderPlot({
  ggplot(data = map) + 
  geom_polygon(aes_string(x = "long", y = "lat", group = "group", fill = input$fill), color = "white") +
  coord_fixed(1.5) + 
  labs(
       x = "Longitude",
       y = "Latitude"
       )
  })
}

shinyApp(ui = ui, server = server)
```

The following show summary statistics for the percentage of visas approved by state, and a summary of the states with the higest and lowest approval rates. 

It's difficult to see a pattern of the state approval rates. Both the maps and the summary show that there are states in both the highest and lowest from each region in the US. One interesting observation is that of the 8 states in which most workers start in the Fall, 5 are among the lost approval rates. 

It's difficult to see a pattern the state approval rates. Both the maps and the summary show that there are states in both the highest and lowest from each region in the US. One interesting observation is that of the 8 states in which most workers start in the Fall, 5 are among the lowest approval rates. However, we were unable to find a reason that fall applications do so poorly. 

```{r state-proportions}
state_props %>%
  summarise(min = min(prop), 
            q1 = quantile(prop, c(0.25)),
            median = median(prop),
            q3 = quantile(prop, c(0.75)),
            max = max(prop),
            mean = mean(prop),
            sd = sd(prop))

state_props %>%
  arrange(desc(prop))

state_props %>%
  arrange(prop)
```

The following tables show the states with the most and fewest applications. There is overlap between the states with the most applications and those with the highest approval rates, and those with the fewest applications and lowest approval rates. This suggests that the number of applications may play a role in the percentage that are approved. 

```{r total-apps}
state_total %>%
  arrange(desc(n))

state_total %>%
  arrange(n)
```


The Federal Office of Management and Budget splits the US into 10 regions, which are more specific than the 4 general regions we used in other parts of the project. In terms of occupation type, 2 of these regions, Region 8 (Colorado, Montana, North Dakota, South Dakota, Utah, Wyoming) and Region 1 (Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island, Vermont) are significant. Every state in these regions, other than Colorado and Connecticut, has "Service" as it's most popular occupation type. There are only 4 other states in the US with this characteristic, and they are scattered through other regions. Generally, the states in this region are somewhat rural and remote, and have low population densities, which may play a role in the occupation type sought by employers for these visas. 

The other 2 states with popular application types other than agriculture are Washington and Iowa. Washington is one of the US's largest producers of lumber, which may help explain why processing is the most popular job type in that state, although Oregon, the US's largest lumber state, does not share this characteristic. Iowa's most popular occupation is Structural Work, which focuses on construction, fabrication, and welding. 

The following tables show summary statistics for the average hourly wage in each state, and the states with the highest and lowest wages. In general, region appears to play a large role in wage. States on the coast tend to pay higher wages, and interior states pay lower ones. This seems to match with coast of living in coastal vs. non coastal states. Regardless, the wages payed to H2B visa workers are generally very low. The median H2B visa worker with a household size of 3 will be near or below the poverty line in 75% of states. Washington pays a significantly high median wage, which may be due to the popular occupation type, processing, in that state. 

```{r state-wages}
state_wages %>%
  summarise(min = min(median_wage), 
            q1 = quantile(median_wage, c(0.25)),
            median = median(median_wage),
            q3 = quantile(median_wage, c(0.75)),
            max = max(median_wage),
            mean = mean(median_wage),
            sd = sd(median_wage))

state_wages %>%
  arrange(desc(median_wage))

state_wages %>%
  arrange(median_wage)
```

Overall the map shows interesting statistics about H2B visas by state. Agricultural work is the most popular in the vast majority of states. Most states also pay very low wages to these workers. Finally, workers in the majority of states start in the Spring. 

```{r - bar graphs}
ui <- fluidPage(
  titlePanel("H2B Visa Data"),
  sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "category", label = "Categories",
                  choices = c("Case Status" = "case_status",
                              "Job Category" = "occupation_category", 
                              "Region" = "region",
                              "Submitted Season" = "submitted_season",
                              "Decision Season" = "decision_season",
                              "Start Season" = "start_season"),
                  selected = "Case Status"),
      selectInput(inputId = "fill", label = "Fill By",
                  choices = c("Region" = "region",
                              "Start Season" = "start_season",
                              "Case Status" = "case_status"),
                  selected = "Region")
    ),
    mainPanel(
      plotOutput(outputId = "bar_graph")
    )
  )

  
)

server <- function(input, output) {
  output$bar_graph <- renderPlot({
    ggplot(data = h2b %>% filter(case_status != "WITHDRAWN" & case_status != "FINAL REVIEW"), mapping = aes_string(x = input$category, fill = input$fill)) +
      geom_bar() + theme_minimal() + 
      labs(x=input$category, y="Count") + 
      coord_flip()
  })
}

shinyApp(ui = ui, server = server)
```

```{r - summary_statistics_bar_graphs}

h2b %>%
   count(case_status)

```


When looking at our bar graphs, we see that most cases are certified. 20,558 cases were approved and 6,657 were partially approved, while only 4204 were denied. 
We are also able to look more into seasonal trends. The most workers submit and receive visa decisions in the winter, and most workers then begin work in the spring and the winter. This makes sense given the workers would start work soon after their visa is approved.

Another interesting observation is that most workers appear to be working in the South. When looking at most frequent job category, we see that most workers are working in the agricultural sector. Since the South produces a large portion of the country's agricultural goods, this makes sense that they would need to bring in seasonal workers to help with harvesting or planting food at different times. 

###Conclusions

If we had more time, we would also include logistic regression to analyze the likelihood of an individual application being approved. This would allow us to set the dependent variable as a categorical one. We would also look at policies that relate to H2-B visas to find if there are any institutional reasons one visa application is more likely to be approved. 


In terms of our dataset, there could be several things pertaining to the reliability and validity of our conclusions. As is the case in most large datasets, we had some NA values that we had to deal with. Since we combined data from several years, different years would sometimes have different columns, so we would get a lot of NA's as a result. We made do with what we had, but missing years of data could mean that changes from year to year were not represented as well. Another hurdle that we had to overcome with our data was the different wage units. Wages were mostly reported as hourly wages, but in some cases they were reported as monthly or yearly wages instead. Since there was no column reporting how many hours each worker worked, there was no way for us to standardize this data. Alas, when examining differences in wages, we chose to mainly foucs on the hourly data since that is what we had the most of to work with. Lastly we had some trouble working with the dates because they were not written in the same format each year. For example, some dates were written as 1/27/08 and others were written as 1-Jan-08. We eventually found a way to make things work, but it took us a while to get things correct. 




